GMSE: an R package for generalised management strategy evaluation

### 2017-11-06 notebook/ms/SI1.Rmd
Finish data structures section

The above array holds all of the information on manager and user actions in the six right-most columns of each array layer. The first seven columns contain information about which resources are affected, and how they are affected. The first column `Act` identifies the type of action being performed; a value of -2 defines a direct action to a resource (e.g., culling of the resource), and a value of -1 defines direct action to a landscape (e.g., increasing yield). Positive values are currently only meaningful for `Manager_Actions`, where a value of 1 defines an action setting a uniform cost of users' direct actions on resources (i.e., costs where `Act = -2` for `User_1_Actions` and `User_2_Actions`). All other values for `Act` are meaningless in GMSE 0.3.1.7, but might be expanded upon in future versions to allow for modification of specific user costs enacted by managers (i.e., managers having different policies for different users) or other users (e.g., users increasing the costs of other users' actions due to conflict or cooperation). Similarly, columns 2-4 refer to resource or landscape types, but only `Type_1 = 1`, `Type_2 = 0`, and `Type_3 = 0 ` are allowed in GMSE v0.3.1.7 (i.e., only one type of resource is permitted), but future versions might allow for different resource types (e.g., `Type_1` might be used to designate species, and `Type_2` and `Type_3` could designate stage or sex). For the rest of this supporting information, we will therefore focus only on rows 1-3 of `ACTION`. Column 5 `Util.` of `ACTION` defines the utility associated with the resource (where `Act = -2`) or landscape (where `Act = -1`). For managers, the target resource abundance set with GMSE argument `manage_target` is found in row 1 (1000 in `ACTION` above); for users, the value in row 1 identifies whether resources are preferred to increase (if positive) or decrease (if negative). Values of column 5 in row 2 similarly identifies whether landscape cell output is preferred by users to increase or decrease (managers do not currently have preferences for landscape output). Of special note is row 3 for `Manager_Actions`, which defines the marginal utility of resources; that is, the adjustment to resource abundance that the manager will attempt to make based on the `manage_target` and the estimated abundance produced by the observation model (in the case of the above, resource abundance is estimated at ca 1065.76, so the manager will set policy in attempt to reduce the population size by ca 65.76 resources). Column 6 `U_land` defines whether or not the utility attached to the resource or landscape output depends on it being on a landscape cell that is owned by the acting user. Related, column 7 `U_loc.` defines whether or not actions can be performed only on a landscape cell that is owned by the acting user. Hence values of columns 6 and 7 are binary, and affected by the `land_ownership` argument in `gmse`. Finally, columns 8-13 correspond to specific actions, either direct (where `Act < 0`) or indirect by setting policy (for row 3 of `Manager_Actions` where `Res_cost = 1`). The last column 13 `None` corresponds with no actions. See [GMSE documentation](https://cran.r-project.org/web/packages/GMSE/GMSE.pdf) for details about the effects of each action.

>>>>

The above array holds all of the information on manager and user actions in the six right-most columns of each array layer. The first seven columns contain information about which resources are affected, and how they are affected. The first column `Act` identifies the type of action being performed; a value of -2 defines a direct action to a resource (e.g., culling of the resource), and a value of -1 defines direct action to a landscape (e.g., increasing yield). Positive values are currently only meaningful for `Manager_Actions`, where a value of 1 defines an action setting a uniform cost of users' direct actions on resources (i.e., costs where `Act = -2` for `User_1_Actions` and `User_2_Actions`). All other values for `Act` are meaningless in GMSE 0.3.1.7, but might be expanded upon in future versions to allow for modification of specific user costs enacted by managers (i.e., managers having different policies for different users) or other users (e.g., users increasing the costs of other users' actions due to conflict or cooperation). Similarly, columns 2-4 refer to resource or landscape types, but only `Type_1 = 1`, `Type_2 = 0`, and `Type_3 = 0 ` are allowed in GMSE v0.3.1.7 (i.e., only one type of resource is permitted), but future versions might allow for different resource types (e.g., `Type_1` might be used to designate species, and `Type_2` and `Type_3` could designate stage or sex). For the rest of this supporting information, we will therefore focus only on rows 1-3 of `ACTION`. Column 5 `Util.` of `ACTION` defines the utility associated with the resource (where `Act = -2`) or landscape (where `Act = -1`). For managers, the target resource abundance set with GMSE argument `manage_target` is found in row 1 (`r samp_ACTION[1,5,1]` in `ACTION` above); for users, the value in row 1 identifies whether resources are preferred to increase (if positive) or decrease (if negative). Values of column 5 in row 2 similarly identifies whether landscape cell output is preferred by users to increase or decrease (managers do not currently have preferences for landscape output). Of special note is row 3 for `Manager_Actions`, which defines the marginal utility of resources; that is, the adjustment to resource abundance that the manager will attempt to make based on the `manage_target` and the estimated abundance produced by the observation model (in the case of the above, resource abundance is estimated at ca `r round((samp_ACTION[1,5,1] + -1*samp_ACTION[3,5,1]),2)`, so the manager will set policy in attempt to change the population size by ca `r round(samp_ACTION[3,5,1],2)` resources). Column 6 `U_land` defines whether or not the utility attached to the resource or landscape output depends on it being on a landscape cell that is owned by the acting user. Related, column 7 `U_loc.` defines whether or not actions can be performed only on a landscape cell that is owned by the acting user. Hence values of columns 6 and 7 are binary, and affected by the `land_ownership` argument in `gmse`. Finally, columns 8-13 correspond to specific actions, either direct (where `Act < 0`) or indirect by setting policy (for row 3 of `Manager_Actions` where `Res_cost = 1`). The last column 13 `None` corresponds with no actions. See [GMSE documentation](https://cran.r-project.org/web/packages/GMSE/GMSE.pdf) for details about the effects of each action.

Constraints on the values that elements in the `ACTION` array can take are defined by a `COST` array (also returned as `manager_array` by `gmse_apply`) of dimensions identical to `ACTION`. All values in `COST` columns 1-7 are set to 10001, one higher than the highest possible budget of either managers or users, so neither can affect resource types or utilities. Columns 8-13 are also set to 10001, except where actions are specifically allowed. Below shows the `COST` array that corresponds to the above `ACTION` array.

```{r, echo = FALSE}
samp_COST <- sim$COST;
dimnames(samp_COST)[[1]] <- c("Resource", "Landscape", "Res_cost", "U1_cost", "U2_cost");
dimnames(samp_COST)[[2]] <- c("Act", "Type_1", "Type_2", "Type_3", "Util.", "U_land", "U_loc.", "Scare", "Cull", "Castrate", "Feed", "Help_off", "None");
dimnames(samp_COST)[[3]] <- c("Manager_Actions", "User_1_Actions", "User_2_Actions");
print(samp_COST);
```

Note that in default GMSE parameters, `culling = TRUE`, but all other actions are false. Hence the `Cull` column 9 is the only column besides column 13 `None` in which cost is less than 10001. Manager's actions in `ACTION` directly affect the cost of users performing one of the five possible actions on resources (columns 8-12). This can be verified in `ACTION` where the manager has set the cost of scaring to `r samp_ACTION[3,9,1]`, and the corresponding `COST` of resource culling (row 1) is `r samp_COST[1,9,2]` for both users. The cost of the manager affecting the cost of user actions is always set to the `minimum_cost` defined in GMSE; here the default `r samp_COST[3,9,1]` is used. This `minimum_cost` also defines cost values for `None`, in which the user or manager does nothing, as might occur if the manager wants to permit culling and therefore does not want to invest any of their `manager_budget` to increasing the cost of culling. Both `ACTION` and `COST` are updated in each time step unless `manage_freq > 1`, in which case `COST` and `Manager_Actions` in `ACTION` are at the frequency defined.
